---
title: "Pipe Dreams Are Made of These"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Dependency
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```
### Read CSV file
```{r}
students <-read_csv("students.csv")
schools <- read_csv("schools.csv")
```
### Preview tibble
```{r}
students %>% head()
```
```{r}
schools %>% head()
```

#### Join two tibbles
```{r message=FALSE, warning=FALSE}
data2 = left_join(students, schools, by=c("school_name"))
data2 %>% head()
```

### Total Number of Schools
```{r}
school_count <- students$school_name %>% 
  unique() %>% 
  length()

school_count
```

### Total Number of Students
```{r}
student_count <-  students %>% nrow()
student_count
```

### Average reading and math scores
```{r}
mean_reading_score <- summarize(students, mean(reading_score))
mean_math_score <- summarize(students, mean(math_score))
```
### Calculate the percentage of students with passing reading scores, i.e. over 70%.
```{r}
percentage_passing_reading <- students %>% 
  filter(reading_score > 70) %>% 
  nrow() * 100 / student_count %>% 
  round(2)

percentage_passing_reading
```
### Calculate the percentage of students with passing math scores, i.e. over 70%.
```{r}
percentage_passing_math <-  students %>% 
  filter(math_score > 70) %>% 
  nrow() * 100 / student_count %>% 
  round(2)

percentage_passing_math
```
### Calculate the overall passing rate, i.e. the average of math and reading passing percentages
```{r}
overall_passing_rate <- (percentage_passing_math + percentage_passing_reading) / 2
overall_passing_rate
```

### Calculate the average math and reading scores by school
```{r}
students %>% 
  group_by(school_name) %>% 
  summarize(avg.reading=mean(reading_score), avg.math=mean(math_score))
```

### Calculate the average math and reading scores by grade level at each school
```{r}
students %>% 
  group_by(school_name, grade) %>% 
  summarize(avg.reading=mean(reading_score), avg.math=mean(math_score))
```
```{r}
total_budget <- schools %>% 
  summarize(sum(budget))
```


### Display data
```{r}
paste("School count: ", school_count)
paste("Student count: ", student_count)
paste("Total budget: ", total_budget)
paste("Average reading score: ", mean_reading_score)
paste("Average math score: ", mean_math_score)
paste("% passing reading: ", percentage_passing_reading)
paste("% passing math: ", percentage_passing_math)
paste("Overall passing rate: ", overall_passing_rate)
```

### Use sapply() to convert data type
```{r}
sapply(school_count, mode)
sapply(school_count, class)
sapply(student_count, mode)
sapply(student_count, class)
sapply(total_budget, mode)
sapply(total_budget, class)
sapply(mean_reading_score, mode)
sapply(mean_reading_score, class)
sapply(mean_math_score, mode)
sapply(mean_math_score, class)
sapply(percentage_passing_math, mode)
sapply(percentage_passing_math, class)
sapply(percentage_passing_reading, mode)
sapply(percentage_passing_reading, class)
sapply(overall_passing_rate, mode)
sapply(overall_passing_rate, class)

# transform(total_budget, numeric = as.double[numeric])
# transform(mean_math_score, numeric = as.double[numeric])
# transform(mean_reading_score, numeric = as.double[numeric])
# transform(percentage_passing_math, numeric = as.double[numeric])
# transform(percentage_passing_reading, numeric = as.double[numeric])
```

### Create Tibble of District Summary
```{r}
# data3 = left_join(students, schools, by=c("school_name"))
# data3 %>% head()
# 
# as_tibble(data3)
# 
# tibble(type,
#        school_name,
#        Avg.Reading.Score = mean_reading_score,
#        Avg.Math.Score = mean_math_score,
#        Total.Students = student_count,
#        Percent.Passing.Reading = percentage_passing_reading,
#        Percent.Passing.Math = percentage_passing_math,
#        Overall.Passing.Rate = overall_passing_rate
# )

# tibble(Total.Schools = school_count,
#        Total.Students= student_count,
#        Total.Budget = total_budget,
#        Avg.Reading = mean_reading_score, 
#        Avg.Math = mean_math_score,
#        Percent.Passing.Reading = percentage_passing_reading,
#        Percent.Passing.Math = percentage_passing_math,
#        Overall.Passing.Rate = overall_passing_rate
# )
# 
# tibble_as(data2)
# type <- group_by(data2, type, school_name, mean_reading_score, mean_math_score, student_count)

```
### Display summary of district-wide data
```{r}
data2 %>% 
  group_by(type, school_name) %>% 
  summarize(avg.reading=mean(reading_score), avg.math=mean(math_score), total.students=sum(student_count), percentage.passing.math=percentage_passing_math, percentage.passing.reading=percentage_passing_reading, percentage.passing.rate=overall_passing_rate)
# select(diamonds, carat, price)
# summarize(data2, sum(school_count))
# summarize(data2, sum(student_count))
# summarize(data2, sum(total_budget))
# summarize(data2, mean(mean_reading_score))
# summarize(data2, mean(mean_math_score))
# summarize(data2, mean(percentage_passing_reading))
# summarize(data2, mean(percentage_passing_math))
# summarize(data2, mean(overall_passing_rate))
```

### Create a per-school summary
```{r}
# data2 %>% 
#   group_by(school_name) %>% 
#   summarize(avg.reading=mean(reading_score), avg.math=mean(math_score), total.students=sum(student_count), percentage.passing.math=percentage_passing_math, percentage.passing.reading=percentage_passing_reading, percentage.passing.rate=overall_passing_rate)

```
```{r}
school_name <- data2 %>% 
  summarize(total.schools=school_count, avg.reading=mean(reading_score), avg.math=mean(math_score), total.students=sum(student_count), percentage.passing.math=percentage_passing_math, percentage.passing.reading=percentage_passing_reading, percentage.passing.rate=overall_passing_rate)
school_name
```